Variational Monte Carlo on Bosonic Systems

### 2018-04-06 project_1/article/article.tex
Spell check the block method

    \subsection{Statistical Analysis}
        If the results of the metropolis sampling were completely uncorrelated,
        it would be enough to compute the standard deviation in a familiar way,
        \begin{equation}
            \sigma = \sqrt{\frac{1}{N}(\ev{E^2_L} - \ev{E_L}^2},
            \label{eq:standard_deviation}
        \end{equation}
        where $N$ is the number of samples, or Monte-Carlo cycles, in the experiment.
        However, it is reasonable to assume that the data we are dealing with in this
        study is liable to suffer from \emph{autocorrelation} and \autoref{eq:standard_deviation}
        does not hold. The prevailing
        definition of autocorrelation in a data stream or signal is correlation
        between a delay of the signal and the original signal. One would be interested to find the
        delay, or lag, in the signal at which the "self-correlation" is highest. We refer to this
        spacing as $d$, and define the following correlation function,
        \begin{equation}
            f_d = \frac{1}{n -d} \sum_{k=1}^{n-d} (x_k - \bar{x}_n)(x_{k+d} - \bar{x}_n).
            \label{eq:correlation_f_d}
        \end{equation}
        The keen reader would have noticed that the function $f_d$ in \autoref{eq:correlation_f_d}
        would be equal to the sample variance for $d=0$. We can now define the \emph{autocorrelation function}
        \begin{equation}
            \kappa_d = \frac{f_d}{\text{Var}(x)},
            \label{eq:autocorrelation_function}
        \end{equation}
        which is equal to $1$ if the data exhibits no autocorrealtions, equating to $d=0$. From the autocorrelation
        function (\ref{eq:autocorrelation_function}) we in turn define the \emph{autocorrelation time},
        \begin{equation}
            \tau = 1 + 2 \sum_{d=1}^{n-1}\kappa_d,
        \end{equation}
        notice that the autocorrelation time is $1$ for a correlation free experiment.

        We are now able to make a correction to the expression for the standard deviation
        improving on \autoref{eq:standard_deviation} by taking correlation into account,
        \begin{equation}
            \sigma =\sqrt{\frac{1 + 2\tau / \Delta t}{N}(\ev{E^2_L} - \ev{E_L}^2)},
        \end{equation}
        where $\Delta t$ is the time between each sample. The main problem at this point is to
        find $\tau$, which we do not know for any given system and it is generally very expensive to
        compute. In order to find a good estimate of $\tau$ we use a procedure called \emph{blocking}.

@@

        The easiest way to do this is by use of the \emph{automatic blocking scheme}, nicely illustrated
        with a flow chart in \autoref{fig:blocking}. The input parameter for the method is an array of the
        local energies, but it can be any array of time series data, which is why it is herein referred to as
        $\mathbf{X}$. Any array of sequential data, where any particular data point depends on the previous
        data point inhibits all attributes of a time series. Our array of local energies is therefor most definitely
        a time series.

        From $\mathbf{X}$ we compute the variance and sample covariance between the series itself and a
        one-point lag of the series. This covariance is usually referred to as the first order autocovariance.
        \begin{align}
            \gamma_i
            &= \frac{1}{n}
            \left(
                \sum_{k=1}^n
                \left((\mathbf{X}_i)_k - (\bar{\mathbf{X}}_i)_k\right)
            \right)
            \nonumber \\
            &\qquad
            \times
            \left(
                \sum_{k=0}^{n-1}
                \left((\mathbf{X}_i)_k - (\bar{\mathbf{X}}_i)_k\right)
            \right).
        \end{align}

        Now we "block"! The next step is to perform a transformation of the our time series $\mathbf{x}_i$, where
        we end up with an array that is half the length of the original time series,
        \begin{equation}
            (\mathbf{X}_i)_k = \frac{1}{2}\left((\mathbf{X}_i)_{2k-1} + (\mathbf{X}_i)_{2k} \right).
        \end{equation}
        This process is were the term "blocking" stems from. There are a number of ways to perform the blocking -
        we could have picked the data points used in the new ($i+1$) array randomly, or more orderly. Here we we have
        picked them sequentially. We have obtained $\frac{n}{2}$ new "primed" stochastic variables. Assuming that the
        reader is playing close attention, it is now needless to say that it is much easier to perform these sequential
        blocking steps if we have an array of length $n= 2^d$, where $d$ is integer. Otherwise, problems would arise
        because one would have arrays of different lengths when the original array is "split in two". The ailment is
        easy to remedy by excluding an observation from the longest resulting array after splitting.

        For each new time series $\mathbf{X}_{i+1}$ a new sample variance and sample covariance is computed.
        The blocking transformation is allowed to continue until we are we run out of data, i.e. the length
        of $\mathbf{X}_i < 2$. In other words, there is only a single number left. We now compute a test statistic,
        \begin{equation}
            M_j = \sum_{k=j}^{d-1} n_k \frac{\left[(n_k - 1)\frac{s^2_k}{n_k^2} y \gamma_k(1)\right] ^2}{s_k^4}.
        \end{equation}
        As luck would have it, this number $M$ is   $\chi^2$-distributed. The following step would be to look up a
        statistical table and find the first $k$, such that $M_k \leq q_{d-k}(1-\alpha)$, where $q$ is the statistical
        limit and $\alpha$ is  the significance level that your hear desires. Finally, the correct sample standard deviation
        is $\frac{s_k^2}{n_k}$. Again, see \autoref{fig:blocking} for a summary of the algorithm.

>>>>

    \subsection{Statistical Analysis}
        If the results of the metropolis sampling were completely uncorrelated,
        it would be enough to compute the standard deviation in a familiar way,
        \begin{equation}
            \sigma = \sqrt{\frac{1}{N}(\ev{E^2_L} - \ev{E_L}^2},
            \label{eq:standard_deviation}
        \end{equation}
        where $N$ is the number of samples, or Monte-Carlo cycles, in the
        experiment.  However, it is reasonable to assume that the data we are
        dealing with in this study is liable to suffer from
        \emph{autocorrelation} and \autoref{eq:standard_deviation} does not
        hold. The prevailing definition of autocorrelation in a data stream or
        signal is correlation between a delay of the signal and the original
        signal. One would be interested to find the delay, or lag, in the signal
        at which the "self-correlation" is highest. We refer to this spacing as
        $d$, and define the following correlation function,
        \begin{equation}
            f_d
            = \frac{1}{n -d} \sum_{k=1}^{n-d}
            (x_k - \bar{x}_n)(x_{k+d} - \bar{x}_n).
            \label{eq:correlation_f_d}
        \end{equation}
        The keen reader would have noticed that the function $f_d$ in
        \autoref{eq:correlation_f_d} would be equal to the sample variance for
        $d=0$. We can now define the \emph{autocorrelation function}
        \begin{equation}
            \kappa_d = \frac{f_d}{\text{Var}(x)},
            \label{eq:autocorrelation_function}
        \end{equation}
        which is equal to $1$ if the data exhibits no autocorrealtions, equating
        to $d=0$. From the autocorrelation function in
        \autoref{eq:autocorrelation_function} we in turn define the
        \emph{autocorrelation time},
        \begin{equation}
            \tau = 1 + 2 \sum_{d=1}^{n-1}\kappa_d.
        \end{equation}
        Notice that the autocorrelation time is $1$ for a correlation free
        experiment.

        We are now able to make a correction to the expression for the standard
        deviation improving on \autoref{eq:standard_deviation} by taking
        correlation into account,
        \begin{equation}
            \sigma =\sqrt{\frac{1 + 2\tau / \Delta t}{N}
            (\ev{E^2_L} - \ev{E_L}^2)},
        \end{equation}
        where $\Delta t$ is the time between each sample. The main problem at
        this point is to find $\tau$, which we do not know for any given system
        and it is generally very expensive to compute. In order to find a good
        estimate of $\tau$ we use a procedure called \emph{blocking}.

@@

            The easiest way to do this is by use of the \emph{automatic blocking
            scheme}, nicely illustrated with a flow chart in
            \autoref{fig:blocking}. The input parameter for the method is an
            array of the local energies, but it can be any array of time series
            data, which is why it is herein referred to as $\mathbf{X}$. Any
            array of sequential data, where any particular data point depends on
            the previous data point inhibits all attributes of a time series.
            Our array of local energies is therefore most definitely a time
            series.

            From $\mathbf{X}$ we compute the variance and sample covariance
            between the series itself and a one-point lag of the series. This
            covariance is usually referred to as the first order autocovariance.
            \begin{align}
                \gamma_i
                &= \frac{1}{n}
                \left(
                    \sum_{k=1}^n
                    \left((\mathbf{X}_i)_k - \bar{\mathbf{X}}\right)
                \right)
                \nonumber \\
                &\qquad
                \times
                \left(
                    \sum_{k=0}^{n-1}
                    \left((\mathbf{X}_i)_k - \bar{\mathbf{X}}\right)
                \right).
            \end{align}
            Note that $i$ represents the iteration in the blocking process
            whereas $k$ is the index into the array $\mathbf{X}_i$.

            Now we "block"! The next step is to perform a transformation of the
            our time series $\mathbf{X}_i$, where we end up with an array that
            is half the length of the original time series,
            \begin{equation}
                (\mathbf{X}_i)_k
                = \frac{1}{2}\left((\mathbf{X}_i)_{2k-1}
                + (\mathbf{X}_i)_{2k} \right).
            \end{equation}
            This process is were the term "blocking" stems from. There are a
            number of ways to perform the blocking - we could have picked the
            data points used in the new ($i+1$) array randomly, or more orderly.
            Here we have picked them sequentially. We have obtained $n/2$ new
            "primed" stochastic variables. Assuming that the reader is playing
            close attention, it is now needless to say that it is much easier to
            perform these sequential blocking steps if we have an array of
            length $n= 2^d$, where $d$ is integer. Otherwise, problems would
            arise because one would have arrays of different lengths when the
            original array is split in two. The ailment is easy to remedy by
            excluding an observation from the longest resulting array after
            splitting.

            For each new time series $\mathbf{X}_{i+1}$ a new sample variance
            and sample covariance is computed.  The blocking transformation is
            allowed to continue until we run out of data, i.e. the length of
            $\mathbf{X}_i < 2$. In other words, there is only a single number
            left.  We now compute a test statistic,
            \begin{equation}
                M_j
                = \sum_{k=j}^{d-1}
                n_k \frac{\left[(n_k - 1)
                \sigma^2_k/n_k^2 + \gamma_k\right] ^2}{\sigma_k^4},
            \end{equation}
            As luck would have it, this number $M$ is   $\chi^2$-distributed.
            The following step would be to look up a statistical table and find
            the first $k$, such that $M_k \leq q_{d-k}(1-\alpha)$, where $q$ is
            the statistical limit and $\alpha$ is your favorite  significance
            level. Finally, the correct sample variance is $\sigma_k^2/n_k$. In
            the rest of the project we use the labeling
            \begin{align}
                \sigma_b^2 \equiv \frac{\sigma_k^2}{n_k},
            \end{align}
            for the correct sample variance found from the blocking method.
            Again, see \autoref{fig:blocking} for a summary of the algorithm.
